---
title: "Analyzing A Single Variant"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(dplyr)
```

## Data Set: A C->G variant

* "0" Reference sequence: C
* "1" Alternate sequence: G

## Reading Data

```{r}
variant <- read.csv("singlevariant.csv", stringsAsFactors = FALSE)
kable(head(variant,n=20))
```


## Visualizing Disease Scores

```{r}
library(ggplot2)
library(ggpubr)
genotypes <- c("CC", "CG", "GG")
variant$Genotype <- factor(genotypes[variant$A12 + 1])
ggplot(variant, aes(x=A12, y=DiseaseScore)) + geom_point() + geom_smooth(method="lm") + theme_pubclean(base_size=20) + xlab("Genotype") + ylab("Disease Score 2")

```

## Box-Whisker Plot of Variant Data

```{r}

# ggplot(variant, aes(x=Genotype, y=DiseaseScore)) + geom_boxplot() + geom_point(color="red", alpha=0.5) + theme_pubclean(base_size=20) + xlab("Genotype") + ylab("Disease Score 2")
ggpubr::ggboxplot(data=variant, x="Genotype", y="DiseaseScore",add = "jitter")
```

## Applying T-Test with `t.test` 

```{r}
scoresCC <- variant[variant$A12==0, "DiseaseScore"]
scoresCGandGG <- variant[variant$A12>0, "DiseaseScore"]
ttestresult <- t.test(scoresCC, scoresCGandGG )
ttestresult
```

## Using R Package `broom`

```{r}
# install.packages("broom")
ttestresult %>% broom::tidy() %>% kable()
```

## Why makes combining CG and GG more sense than comining CC and CG

* Reference sequence: C
* Alternate sequence: G

## Linear Model

```{r}
linmodel <- lm(DiseaseScore ~ A12, data=variant)
summary(linmodel)
```

## Linear Model with Broom

```{r}
linmodel %>% broom::tidy() %>% kable()
```


## Visualizing Age versus DiseaseScore

```{r}
ggplot(variant, aes(x=Age, y=DiseaseScore)) + geom_point() + geom_smooth(method="lm")
```

## Linear Model DiseaseScore as function of Age

```{r}
lm(DiseaseScore ~ Age, data=variant) %>% broom::tidy() %>% kable()
```

## Visuializing Age vs A12

```{r}
ggplot(variant, aes(x=Genotype, y=Age)) + geom_boxplot()
```

## Linear Model of Copy number versus Disease Score 2 Adjusted for Age

```{r}
linmodel <- lm(DiseaseScore ~ A12 + Age, data=variant)
linmodel %>% broom::tidy() %>% kable()
```
