---
title: "Batch Effects and Confounding"
author: "Eckart Bindewald"
date: "1/28/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

## Batch Effects

"The batch effect represents the systematic technical differences when samples are processed and measured in different batches and which are unrelated to any biological variation [...]"

Leek JT, Johnson WE, Parker HS, Jaffe AE, Storey JD. The sva package for
removing batch effects and other unwanted variation in high-throughput
experiments. Bioinformatics. 2012 

## Examples of Batch Effects

* Laboratory conditions
* Choice of reagent lot or batch
* Personnel differences
* Time of day when the experiment was conducted
* Atmospheric ozone levels
* Instruments used to conduct the experiment

## Simpson's Paradox

* Data from UC Berkely Admissions Office (1973)
* Lhe chance of being admitted among the applicants was higher for men then women
* Led to law suit claiming discrimination

## Berkely Admission Data in R

3D Dataset of [Admitted/Rejected,Male/Female,Department A-F]

```{r}
UCBAdmissions
```

## Summing over Departments

```{r}
apply(UCBAdmissions, c(1, 2), sum)
```

## Exercise using Dataset UCBAdmissions

* How many men applied in 1973?
* How many women applied in 1973?
* How many men (women) were accepted?
* Compute the chance of being admitted for male and for female applicants using dataset UCBAdmissions

## Mosaic Plots

```{r}
mosaicplot(apply(UCBAdmissions, c(1, 2), sum),
           main = "Student admissions at UC Berkeley")
```

## Exercise

* Create mosaic plots for the individual departments A,B,...,F
* If needed, you can obtain help in R via ?UCBAdmissions
* Is there a gender-bias trend that is consistent for all departments?
* What could be an explanation leading to the observed apparent bias in the aggregated data?

## Confounding

Definition: 

* Correlation is not causation
* Sometimes X ~ Y (correlation) because causation due to other variable Z: X~Z and Y~Z
* We say Z is a confounding variable

## Downloading Data

```{r}
if (!require(sva)) {
  BiocManager::install("sva")
  require(sva)
}
arafile <- "arabidopsis.RDS"
url <- paste0("https://github.com/PacktPublishing/R-Bioinformatics-Cookbook/blob/master/datasets/Chapter01/", arafile,"?raw=true")
if (!file.exists(arafile)) {
  download.file(url, arafile)
}
arab <- readRDS(arafile)
```

## Filtering out Rows

```{r}
keep <- apply(arab, 1, function(x) { length(x[x>3])>=2 } )
arab_filtered <- arab[keep,]
```

## Creating the Initial Design

```{r}
groups <- as.factor(rep(c("mock", "hrcc"), each=3))
```

## Test and Null Models

```{r}
test_model <- model.matrix(~groups)
null_model <- test_model[,1]
svar <- svaseq(arab_filtered, test_model, null_model, n.sv=1)
```

## Extract the surrogate variables

```{r}
design2 <- cbind(test_model, svar$sv)
design2
```

