---
title: "Bio3D R pagage"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## bio3d R package

```{r}
# install.packages("bio3d)
library(bio3d)
```


```{r}
pdbfile = system.file("examples/1hel.pdb", package="bio3d") 
pdb <- read.pdb( pdbfile)
x = read.pdb("4dfr.pdb")
names(x)
```

## Data structure "atom"

```{r}
x[["atom"]]
```


## Data structure "xyz"

```{r}
x[["xyz"]]
```

## Selecting Atoms with "atom.select"

```{r}
# Select protein atoms of chain A
chaina <- atom.select(pdb, "protein", chain="A")
```

## Selecting Atoms 

```{r}
# Select all C-alpha atoms with residues numbers between 43 and 54
sele <- atom.select(pdb, "calpha", resno=43:54, verbose=TRUE)
# Access the PDB data with the selection indices
print( pdb$atom[ sele$atom, "resid" ] )
```

## Create PDB Data Structure Based on Selection

```{r}
# Trim PDB to selection
ca.pdb <- trim.pdb(pdb, sele)
ca.pdb
```

## Exercise

* Select backbone atoms
* Select residue numbers 1-5
* Select atom number 1-30

## Solution 1: Use all selectors at once

```{r}
backbone <- atom.select(pdb,"backbone", resno=1:5, eleno=1:30)
head(pdb$atom[backbone$atom, ], n=100) 
pdb_backbone <- trim.pdb(pdb, backbone)
```

## Solution 1: Use all selectors at once

```{r}
backbone_selection <- atom.select(pdb,"backbone")
resno_selection<- atom.select(pdb,"protein", resno=1:5)
eleno_selection <- atom.select(pdb,"protein", eleno=1:30)
combo_selection <- combine.select(backbone_selection, resno_selection, eleno_selection, operator = "AND")

head(pdb$atom[backbone$atom, ], n=100) 
pdb_backbone_combo <- trim.pdb(pdb, combo_selection)
pdb_backbone_combo
```

## Exercise:

* Try combine.select but use operator="OR"

```{r}
pdb_backbone_or <- trim.pdb(pdb, combo_selection_or)
# write.pdb(pdb_backbone_or, file="~/pdb_backbone_or.pdb")
```

## Visualize 

`http://jena3d.leibniz-fli.de/`



## Aligning PDB Structures

```{r}
root = system.file(package="bfxapps2")
pdbfile = file.path(root, "extdata","pdb", "1xwc.pdb")
pdbfile2 = file.path(root, "extdata","pdb", "3trx.pdb")
stopifnot(file.exists(pdbfile))
stopifnot(file.exists(pdbfile2))
a <- read.pdb(pdbfile)
b <- read.pdb(pdbfile2)
stopifnot(file.exists(pdbfile))
# download.file(pdbfile, "1xwc.pdb")
```

## Aligning PDB Structures

```{r}
# aln <- pdbaln(list("1xwc"=a, "3trx"=b))
# aln
```





