---
title: "Machine Learning with Random Forests"
output:
  powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

## Preliminaries

```{r}
library(dplyr)
library(knitr)
library(randomForest)

train_rows <- sample(nrow(iris), 0.8 * nrow(iris), replace=FALSE)
train_set <- iris[train_rows,]
test_set <- iris[-train_rows,]

```

## Train the model

```{r}
model <- randomForest(Species ~ ., data=train_set, mtry = 2,
                       importance=TRUE)
model
```

## Prediction Accuracy

```{r}
test_set_predictions <- predict(model, test_set, type="class")
caret::confusionMatrix(test_set_predictions, test_set$Species)
```

## Feature Importance

```{r}
varImpPlot(model)
```

## Exercises

1. Compute the accuracy of the prediction with respect to the test set based
on the confusion matrix.
The accurracy is the ratio of correctly predicted cases divided by the total number of predicted cases.

2. Change the evaluation of the model (function predict(...)) such that instead
of the test set the training set is used (change test_set to train_set).
What happens to the prediction accuracy? Why is this not "allowed"?

3. Create 4 models that depend on only 3 parameters and store them in different variables:
a) Sepal.Length, Sepal.Width, Petal.Length
b) Sepal.Length, Sepal.Width, Petal.Width
c) Sepal.Length, Petal.Length, Petal.Width
d) Sepal.Width, Petal.Length, Petal.Width
Compute the accuracy for each model. Which of the 4 models has the 
highest accuracy

4) Improve the Markdown output of the confusion matrix by applying functions
as.matrix , as.data.frame and knitr::kable


