---
title: "Multiple Sequence Alignments"
author: "Eckart Bindewald"
date: "2/11/2020"
output:
  ioslides_presentation:
    smaller: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## R Package "muscle"

* Performs multiple sequence alignments on DNA or protein sequences
* Similar accuracy to program ClustalW but faster
* Edgar RC (2004). “MUSCLE: multiple sequence alignment with high accuracy and high throughput.” Nucleic Acids Research

```{r}
library(muscle)
```


## Prepared Unaligned Sequences

```{r}
umax
```

## Aligning Sequences

```
result <- muscle(umax,diags=TRUE)
```

## R Class DNAMultipleAlignment

* Stores Results of Multiple Sequence Alignments
* Core class in Biostrings R package
* <https://www.bioconductor.org/packages/devel/bioc/vignettes/Biostrings/inst/doc/MultipleAlignments.pdf>

## Aligned Sequences

```{r include=FALSE}
result <- muscle(umax,diags=TRUE)
```

```{r}
result
```

## Alphabet Frequency

```{r}
alphabetFrequency(result)
```

## Consensus Matrix

```{r}
consensusMatrix(result)[,1:12]
```

## Consensus Matrix

```{r}
consensusMatrix(result)[1:4,1:12]
```

## Tree Similarity

```{r}
sdist <- stringDist(as(result,"DNAStringSet"), method="hamming")
clust <- hclust(sdist, method = "single")
plot(clust)
```
