---
title: "Sequence Alignments using Clustal"
author: "Eckart Bindewald"
date: "4/14/2020"
output:
  powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## R Library msa

```{r}
library(msa)
```

## Download Data

```{r}
remote_folder="https://raw.githubusercontent.com/PacktPublishing/R-Bioinformatics-Cookbook/master/datasets/Chapter03"
fa_file = "hglobin.fa"
if (!file.exists(fa_file)) {
  download.file(file.path(remote_folder,fa_file),fa_file)
}
stopifnot(file.exists(fa_file))
```

## Read Alignment

```{r}
library(Biostrings)
seqs = readAAStringSet(fa_file)
```

## Perform Sequence Alignment

```{r}
alignment = msa(seqs, method="ClustalOmega")
alignment
```

## "Pretty Print" of Sequence Alignment

```{r}
msaPrettyPrint(alignment, output="pdf",file="hglobin_msa.pdf",askForOverwrite = FALSE)
```


## Show The Create Print

```{r}
knitr::include_graphics("hglobin_msa.pdf")
```


## "msa" as frontend for "Clustal","ClustalOmega" and "Muscle" Algorithms

```{r}
alignment_muscle = msa(seqs, method="Muscle")
alignment_muscle
```

## Summary

* The library "msa" has the R function "msa" for computing multiple sequence alignments
* Function "msa" provides uniform frontend for different alignment programs
* Implemented algorithms: Clustal, ClustalOmega, Muscle
* Function msaPrettePrint for publication-quality output



