---
title: "Analyzing Single Variant"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(dplyr)
```

## Inspect Data

```{r}
library(bfxapps2)
data(singlevariant)
variant = singlevariant
kable(head(variant))
variant$Age = 20*rnorm(nrow(variant)) + 35
variant$DiseaseScore2 = variant$DiseaseScore + variant$Age * 0.1 * runif(nrow(variant))
```

## Compute Number of Variant Allels

* This variant describes an C->G substitution. "0" stands for C, "1" stands for G at that position
* We can have allele 1 or 2 or both be reference or alternative sequence
* Assumption that phenotype is more pronounced the higher the number of variant alleles 
* Number of variant alleles?

## Add Column Corresponsing to Number of Variant Alleleles


```{r}
variant$A12 = variant$A1 + variant$A2
```

## Visualize Data: Scatter Plot

```{r}
library(ggplot2)
ggplot(variant, aes(x=A12, y=DiseaseScore2)) + geom_point()
```

## Visualize Data: Box-Plot

```{r}
library(ggplot2)
ggplot(variant, aes(x=factor(A12), y=DiseaseScore2)) + geom_boxplot()
```


## Your Job: Improve Output!

* Try to change this plot to make it closer to publication-quality

```{r}
library(ggplot2)
ggplot(variant, aes(x=factor(A12), y=DiseaseScore2)) + geom_point()
```

## Add Linear Regression in Plot

```{r}
ggplot(variant, aes(x=factor(A12), y=DiseaseScore2)) + geom_point() + geom_smooth(method="lm")
```

## Linear Regression Data

```{r}
lm(DiseaseScore2 ~ A12, data = variant) %>% broom::tidy() %>% kable()
```

## Dependence on Age?

```{r}
ggplot(variant, aes(x=Age, y=DiseaseScore2)) + geom_point()
```

## Find out if Dependence on Age is Statistically Significant

```{r}
ggplot(variant, aes(x=Age, y=DiseaseScore2)) + geom_point()
```

## Building Adjusted Models

```{r}
lm(DiseaseScore2 ~ A12 + Age, data = variant) %>% broom::tidy() %>% kable()
```



