---
title: "Translation"
author: "Eckart Bindewald"
date: "4/28/2020"
output:
  powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Preliminaries

```{r}
library(Biostrings)
```

## Translation Protein Sequences - Basic Example

```{r}
dna1 <- DNAString("TTGATATGGCCCTTATAA")
translate(dna1)
```


## Alternative Genetic Codes

* Mitochondria and archea can have alternative genetic codes

```{r}
SGC1 <- getGeneticCode("SGC1")  # Vertebrate Mitochondrial code
translate(dna1, genetic.code=SGC1)
```

## Translating an Open Reading Frame

```{r}
file <- system.file("extdata", "someORF.fa", package="Biostrings")
x <- readDNAStringSet(file)
x
```

## The first and last 1000 nucleotides are not part of the ORFs:

```{r}
x <- DNAStringSet(x, start=1001, end=-1001)
```

* Before calling translate() on an ORF, we need to mask the introns
* Use database (in this case SGD, http://www.yeastgenome.org/).
* According to SGD, the 1st ORF (YAL001C) has an intron at 71..160
* see <http://db.yeastgenome.org/cgi-bin/locus.pl?locus=YAL001C>

```{r}
y1 <- x[[1]]
mask1 <- Mask(length(y1), start=71, end=160)
masks(y1) <- mask1
y1
translate(y1)
```

## Codons:

```{r}
codons(y1)
which(width(codons(y1)) != 3)
codons(y1)[20:28]
```


## Translation of negative Strand Transcripts

Translation on the '-' strand:

```{r}
dna3 <- DNAStringSet(c("ATC", "GCTG", "CGACT"))
translate(reverseComplement(dna3))
```

## What if we do not know the reading frame?

## Translate sequences on both '+' and '-' strand across all 
## possible reading frames (i.e., codon position 1, 2 or 3):
## First create a DNAStringSet of '+' and '-' strand sequences, 
## removing the nucleotides prior to the reading frame start position.

```{r}
dna3_subseqs <- lapply(1:3, function(pos) 
    subseq(c(dna3, reverseComplement(dna3)), start=pos))
## Translation of 'dna3_subseqs' produces a list of length 3, each with
## 6 elements (3 '+' strand results followed by 3 '-' strand results).
lapply(dna3_subseqs, translate)
```

