---
title: "Tree With Branch Lengths"
author: "Eckart Bindewald"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggtree)
library(treeio)
```

With ggtree, it is straight-forward, to label tree tip nodes (with `geom_tiplab`).
But how can one label tree branch lengths? This is not so obvious and currently
restricted to certain tree layouts.

In non-circular (default) ggtree layout, the x-coordinate coincides with the 
need coordinate of the branch labeling. This can be used with the `geom_nodelab`
function as shown in this example:

```{r}
file <- system.file("extdata/BEAST", "beast_mcc.tree", package="treeio")
beast <- read.beast(file)
beast
ggtree(beast, aes(color=rate))  +
    geom_nodelab(aes(x=branch, label=round(length,1)), vjust=-.5, size=3) +
#      geom_nodelab(aes(x=branch, label=round(posterior, 2)), vjust=-.5, size=3) +
    geom_tiplab() +
    scale_color_continuous(low="darkgreen", high="red") +
    theme(legend.position=c(.1, .8))
# + geom_range(range='length_0.95_HPD', color='red', alpha=.6, size=2) 
```


## Credits

This page has been adapted from a web page maintained by the author and maintainer of the `ggtree`
R package (Guangchuang Yu): 

<https://yulab-smu.top/treedata-book/chapter5.html>
